---
title: "qPCR QAQC of results"
author: "Mariah Kachmar"
date: "2025-07-09"
output: html_document
Description: The purpose of this script is to import, clean up, and summarize Triplex qPCR data for the detection of Dermo, MSX, and SSO to determine prevalence and infection intensity for the LISS Oyster Health Project's monthly sampling at Ash Creek and Fence Creek intertidal sites in Connecticut and Goldstar beach and Laurel Hollow subtidal sites on Long Island, NY. Infection intensity will be reported as copies/ng. This data will then be cross validated with RFTM and Histology data. 

Last updated:9/3/26 by Mariah kachmar

- working on inputting files & a QAQC routine

---

#Set working directory
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/mariah.kachmar/documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance') 
#knitr::opts_knit$set(root.dir = 'C:/Users/kyra.lenderman/documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance') 
#knitr::opts_knit$set(root.dir = 'C:/Users/sarina.dery/documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance') 
```

#run packages
```{r}
library("ggplot2")
library("readxl")
library("plyr")                                     
library("dplyr")
library("Rmisc")
library("readr")  
library("lubridate")
library("purrr")
library(reshape2)
library(stringr)
library("RColorBrewer")
library("stringr")
library("tidyverse")
```


#### This code chunk merges all .csv files within the qubit folder into one data fram and outputs the full dataset into a .csv master file. This allows us to download the raw data as a .csv, add it to the repository folder, and create the master data file without copying and pasting data in excel.
# reading in .csv files from local folder
```{r}


data_all <- list.files(path = "Lab_Data_qPCR/raw_data",pattern = "\\.csv$",full.names = TRUE) %>%
  # Read each CSV file
  lapply(function(file_path) {
    read_csv(file_path,skip = 19,show_col_types = FALSE) %>%
      # Add filename so we can identify the source file
      mutate(File_Name = basename(file_path))}) %>% 
  # Combine all CSV files
  bind_rows() %>%
  # Remove rows where there is no sample
  filter(!is.na(Content), Content != "") %>%
  # Keep only the columns needed
  dplyr::select(Fluor,Content, Cq,`Starting Quantity (SQ)`,`Cq Mean`,`Cq Std. Dev`,`SQ Std. Dev`,File_Name) %>%
  # Rename columns to be more user-friendly
  dplyr::rename(
    Target_dye = Fluor,
    Sample_ID = Content,
    Concentration = `Starting Quantity (SQ)`,
    St_dev = `SQ Std. Dev`,
    Cq_st_dev = `Cq Std. Dev`,
    Cq_mean = `Cq Mean`) %>%
  # Add filename to standard IDs
  mutate(Sample_ID = if_else( str_detect(Sample_ID, "^Std-0[1-8]$"),paste0( tools::file_path_sans_ext(File_Name), "_",Sample_ID),
      Sample_ID)) %>%
  # Extract information from Sample_ID
  mutate(
    # First 4 characters = month/year
    Month_year = str_sub(Sample_ID, 1, 4),
    # Extract site name between month/year and replicate number
    Site = str_extract(Sample_ID, "(?<=\\d{4})[A-Za-z]+(?=_)"),
    # Extract replicate number after underscore
    Replicate = str_extract( Sample_ID, "(?<=_)\\d+"),
    # Convert month/year to a real date
    Date = my(Month_year))

# View the final dataset
data_all

write.csv(data_all, "Master_files\\qPCR_all_data.csv", row.names=FALSE)
```

# identify when 10^1 doesn't amplify and exclude samples that fall within that range for downstream analysis. If this becomes a consistent problem, or a high number of samples are falling within that range, then the samples will need to be rerun. 
```{r}
# Check Std-08 on each plate/file
data_all <- data_all %>%
  mutate(
    Standard = case_when(
      str_detect(Sample_ID, regex("_Std-01$", ignore_case = TRUE)) ~ "10^8",
      str_detect(Sample_ID, regex("_Std-02$", ignore_case = TRUE)) ~ "10^7",
      str_detect(Sample_ID, regex("_Std-03$", ignore_case = TRUE)) ~ "10^6",
      str_detect(Sample_ID, regex("_Std-04$", ignore_case = TRUE)) ~ "10^5",
      str_detect(Sample_ID, regex("_Std-05$", ignore_case = TRUE)) ~ "10^4",
      str_detect(Sample_ID, regex("_Std-06$", ignore_case = TRUE)) ~ "10^3",
      str_detect(Sample_ID, regex("_Std-07$", ignore_case = TRUE)) ~ "10^2",
      str_detect(Sample_ID, regex("_Std-08$", ignore_case = TRUE)) ~ "10^1",
      TRUE ~ NA_character_
    )
  )

data_all

std08_check <- data_all %>%
  dplyr::filter(Standard == "10^1") %>%
  dplyr::group_by(File_Name, Target_dye) %>%
  dplyr::summarise(
    n_Std08 = length(Cq),
    n_amplified = sum(!is.na(Cq)),
    Std08_amplified = any(!is.na(Cq)),
    .groups = "drop"
  )

std08_check

failed_10_1_files <- std08_check %>%
  filter(Std08_amplified == FALSE) %>%
  select(File_Name, Target_dye)
failed_10_1_files

std10_2 <- data_all %>%
  dplyr::filter(
    str_detect(
      Sample_ID,
      regex("_Std-07$", ignore_case = TRUE)
    )
  ) %>%
  dplyr::group_by(File_Name, Target_dye) %>%
  dplyr::summarise(
    Cq_10_2 = first(Cq),
    .groups = "drop"
  )

std10_2


unknowns_failed_10_1 <- data_all %>%
  # Keep only files where 10^1 failed
  inner_join(failed_10_1_files,by = c("File_Name", "Target_dye") ) %>%
  # Remove standards
  filter(!str_detect(Sample_ID, regex("_Std-0[1-8]$", ignore_case = TRUE) ) ) %>%
  # Add the 10^2 standard Cq for that plate
  left_join(std10_2, by = c("File_Name", "Target_dye") ) %>%
  # Identify samples that amplified after the 10^2 standard
  mutate(
    Potential_10_1_to_10_2 = case_when(
      # No Cq = did not amplify
      is.na(Cq) | is.nan(Cq) ~ FALSE,
      # Cq later than 10^2 standard
      Cq >= Cq_10_2 ~ TRUE,TRUE ~ FALSE))

potential_low_copy <- unknowns_failed_10_1 %>%filter(Potential_10_1_to_10_2 == TRUE)

potential_low_copy

#clean the data
data_all_clean <- data_all %>%anti_join(potential_low_copy %>% select(File_Name, Target_dye, Sample_ID),by = c("File_Name", "Target_dye", "Sample_ID"))

data_all_clean

#check to see if samples were removed
nrow(data_all)
nrow(data_all_clean)

nrow(data_all) - nrow(data_all_clean)

#file to keep track of samples excluded
excluded_low_copy <- potential_low_copy %>%
  select(File_Name,Target_dye,Sample_ID,Cq,Cq_10_2) %>%
  mutate(Exclusion_reason = "Potential 10^1–10^2 copy sample; 10^1 standard failed")
excluded_low_copy
```


```{r}
#Percentage of standard deviations >0.5 for qPCR Cq values
high_cq_sd <- data_all %>% filter(Cq_st_dev > 0.5) %>% nrow() #creates number of rows that have sd # >0.5
high_cq_sd_list <- data_all %>% filter(Cq_st_dev > 0.5) #creates list of which samples have high sd
percentage_highsd <- (high_cq_sd/nrow(data_all)) * 100 # determining percentage of samples have high standard deviation
percentage_highsd #outputs percentage


#Percentage of standard deviations >0.5 for qPCR technical replicate values
high_sd <- data_all %>% filter(St_dev > 0.5) %>% nrow() #creates number of rows that have sd # >0.5
high_sd_list <- data_all %>% filter(St_dev > 0.5) #creates list of which samples have high sd
percentage_highsd <- (high_sd/nrow(data_all)) * 100 # determining percentage of samples have high standard deviation
percentage_highsd #outputs percentage

#Identify which samples have a standard deviation > 0.5 for qPCR technical replicate values
high_sd_list <- data_all %>% filter(St_dev >0.05) %>% arrange(desc(St_dev))
high_sd_list

```

```{r}
#Lets create means of our sample duplicates 

data_summary <- data_all_clean %>%
  dplyr::mutate(Concentration = as.numeric(Concentration)) %>%
  dplyr::group_by(Target_dye,Sample_ID, Site, Date) %>%
  dplyr::summarise(
    Copies_per_2ul = mean(Concentration, na.rm = TRUE),
    Copies_per_2ul_sd = sd(Concentration, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

data_summary


#now we need copies per ul for reporting

data_summary <- data_summary %>%
  mutate(
    Copies_per_uL =  Copies_per_2ul / 2,
    Copies_per_uL_sd =  Copies_per_2ul_sd / 2
  )
  
data_summary



```

```{r}

#merge Qubit data

#bring in data
data_quibit <- read.csv("Lab_Data_DNAQuantification/output/qubit_summary.csv")
data_quibit


merged_data <- merge(data_summary, data_quibit, by = "Sample_ID")
merged_data

```

```{r}

#Lets complete those back calculations

data_summary <- merged_data %>%
  mutate(
    Ratio_copies_per_ng =  Copies_per_uL / mean_concentration_ng.uL,
    Copies_ul_ng.ul =  Ratio_copies_per_ng * 33.33, #this will be the final concentration per 1 ul, roger williams uses this concentration (what we should report for scoring?)
    Copies_2ul_final = Copies_ul_ng.ul *2 #this is the final concentration per reaction
  )
  
data_summary

#add a year column 
data_summary <- data_summary %>%
  mutate(Year = year(Date))

```

```{r}

#lets work on some code that bins our copies/ul into the Mackin's scale. This is the same scale that Roger Williams Univ. is using currently and may be subject to change based on our data. This is preliminary to test code!!

#Let's parse our dermo data (Target_dye = FAM)

dermo_qPCR_summary <- data_summary %>% filter(Target_dye == 'FAM')
dermo_qPCR_summary

qpcr_mackin_scores <- dermo_qPCR_summary %>%
  mutate(mackin_score = case_when(
    Copies_ul_ng.ul < 10 ~ 0,
    Copies_ul_ng.ul >= 10 & Copies_ul_ng.ul < 100 ~ 0.5, #light
    Copies_ul_ng.ul >= 100 & Copies_ul_ng.ul < 2000 ~ 1, #light
    Copies_ul_ng.ul >= 2000 & Copies_ul_ng.ul < 10000 ~ 2, #moderate
    Copies_ul_ng.ul >= 10000 & Copies_ul_ng.ul < 100000 ~ 3, #moderate
    Copies_ul_ng.ul >= 100000 & Copies_ul_ng.ul < 1000000 ~ 4, #heavy
    Copies_ul_ng.ul >= 1000000 ~ 5, #heavy
    TRUE ~ 0
  ))

qpcr_mackin_scores

#Lets do the same for MSX (Target_dye = HEX)


MSX_qPCR_summary <- data_summary %>% filter(Target_dye == 'HEX')
MSX_qPCR_summary

MSX_mackin_scores <- MSX_qPCR_summary %>%
  mutate(mackin_score = case_when(
    Copies_ul_ng.ul < 10 ~ 0,
    Copies_ul_ng.ul >= 10 & Copies_ul_ng.ul < 10000 ~ 1,
    Copies_ul_ng.ul >= 10000 & Copies_ul_ng.ul < 1000000 ~ 2,
    Copies_ul_ng.ul >= 1000000  ~ 3,
    TRUE ~ 0
  ))

MSX_mackin_scores

```

```{r}

#Now for some prevalence and intensity calculations

#weighted prevalence 

#dermo
d_wp_summary <- qpcr_mackin_scores %>%
  dplyr::group_by(Site, Date, Year) %>% 
  dplyr::summarize( Weighted_prevalence = mean(mackin_score))
d_wp_summary

#dermo graph
ggplot(data=d_wp_summary, aes(x=Date, y=Weighted_prevalence, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Weighted Prevalence of Dermo infection (qPCR)", x ="Date", y = "Weighted Prevalence Score")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "fixed", ncol=1,)+ scale_x_date(
  date_breaks = "1 month",
  date_labels = "%b/%Y")#scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))

#MSX
msx_wp_summary<- MSX_mackin_scores %>%
  dplyr::group_by(Site, Date, Year) %>% 
  dplyr::summarize( Weighted_prevalence = mean(mackin_score))

msx_wp_summary

#MSX graph

ggplot(data=msx_wp_summary, aes(x=Date, y=Weighted_prevalence, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Weighted Prevalence of MSX infection (qPCR)", x ="Date", y = "Weighted Prevalence Score")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "fixed", ncol=1,)+ scale_x_date(
  date_breaks = "1 month",
  date_labels = "%b/%Y")#scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))

```

# Prevalence calculations for all sites
# Prevalence = # infected/ total # sampled
```{r}

## Prevalence T/F for DERMO
qpcr_mackin_scores<-qpcr_mackin_scores%>% filter(!is.na(mackin_score))
qpcr_mackin_scores

qpcr_mackin_scores$prevalent = ifelse(qpcr_mackin_scores$mackin_score == 0, '0',
  ifelse(qpcr_mackin_scores$mackin_score > 0, '1', NA))
  
qpcr_mackin_scores

# YEARLY PREVALENCE %
present <- nrow(qpcr_mackin_scores %>% filter(prevalent == '1'))
absent <- nrow(qpcr_mackin_scores %>% filter(prevalent == '0'))

prevalence_all <- present/(present+absent)
prevalence_all


str(qpcr_mackin_scores$prevalent)
qpcr_mackin_scores$prevalent <- as.numeric(qpcr_mackin_scores$prevalent)

d_prevalence_summary <- qpcr_mackin_scores %>%
  dplyr::group_by(Site, Year, Date) %>% 
  dplyr::summarize( prevalence_precentage= mean(prevalent)* 100)
d_prevalence_summary


#Graph DERMO

ggplot(data=d_prevalence_summary, aes(x=Date, y=prevalence_precentage, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Prevalence of Dermo infection (qPCR)", x ="Date", y = "Dermo Prevlaence %")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "free", ncol=1) + scale_x_date(
  date_breaks = "1 month",
  date_labels = "%b/%Y")#+ scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))


## Prevalence T/F for MSX

MSX_mackin_scores<-MSX_mackin_scores%>% filter(!is.na(mackin_score))
MSX_mackin_scores

MSX_mackin_scores$prevalent = ifelse(MSX_mackin_scores$mackin_score == 0, '0',
  ifelse(MSX_mackin_scores$mackin_score > 0, '1', NA))
  
MSX_mackin_scores

# YEARLY PREVALENCE %
present <- nrow(MSX_mackin_scores %>% filter(prevalent == '1'))
absent <- nrow(MSX_mackin_scores %>% filter(prevalent == '0'))

prevalence_all <- present/(present+absent)
prevalence_all


str(MSX_mackin_scores$prevalent)
MSX_mackin_scores$prevalent <- as.numeric(MSX_mackin_scores$prevalent)

MSX_prevalence_summary <- MSX_mackin_scores %>%
  dplyr::group_by(Site, Year, Date) %>% 
  dplyr::summarize( prevalence_precentage= mean(prevalent)* 100)
MSX_prevalence_summary

#Graph MSX

ggplot(data=MSX_prevalence_summary, aes(x=Date, y=prevalence_precentage, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Prevalence of MSX infection (qPCR)", x ="Date", y = "MSX Prevalence %")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "free", ncol=1)+ scale_x_date(
  date_breaks = "1 month",
  date_labels = "%b/%Y")#scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))


```

#Lets compare the RFTM scores to the qPCR. qPCR is more sensitive so we may see slightly inflated scores but I think it is good to check in terms of postive and negative results. RFTM we are also only taking mantle tissue vs qPCR we are taking GGR. What I am looking for is if qPCR is lower than RFTM or is negative and RFTM is positive. 
```{r}
#pull in RFTM data
RFTM_all <- read.csv("Master_files/RFTM_all_data.csv")
RFTM_all <- RFTM_all %>% dplyr::rename("Sample_ID" = "lab_id")

#Merge dermo qPCR data

rftm_v_qpcr <- merge(RFTM_all, qpcr_mackin_scores, by = "Sample_ID")
rftm_v_qpcr <- rftm_v_qpcr %>% select(Sample_ID, site, rftm_final, mackin_score , Copies_ul_ng.ul)
rftm_v_qpcr

rftm_v_qpcr$results = ifelse(rftm_v_qpcr$mackin_score < rftm_v_qpcr$rftm_final, 'FALSE',
  ifelse(rftm_v_qpcr$mackin_score > rftm_v_qpcr$rftm_final, 'TRUE',
  ifelse(rftm_v_qpcr$mackin_score == rftm_v_qpcr$rftm_final, 'TRUE', NA)))
  
rftm_v_qpcr

rftm_v_qpcr %>%
  dplyr::filter(
    !is.na(rftm_final),
    !is.na(mackin_score)
  ) %>%
  dplyr::summarise(
    n = length(rftm_final),
    qPCR_positive_RFTM_zero = sum(
      rftm_final == 0 & mackin_score > 0
    ),
    percent = sum(
      rftm_final == 0 & mackin_score > 0
    ) / length(rftm_final) * 100
  )

rftm_v_qpcr %>%
  dplyr::filter(
    rftm_final == 0,
    mackin_score > 0
  ) %>%
  dplyr::count(mackin_score, name = "Frequency")

FALSE_qpcr<- rftm_v_qpcr %>% filter( results == 'FALSE') %>%
  mutate(
    rerun_assay = NA_character_,
    rerun_date = as.Date(NA),
    rerun_result = NA_real_,
    rerun_mackin_score = NA_real_,
    rerun_status = "Pending"
  ) #adding columns we can fill in later on once we rerun the samples to identify if these results are true
FALSE_qpcr

write.csv(FALSE_qpcr, "Lab_Data_qPCR\\outputs/qPCR_sample_rerun_list.csv", row.names=FALSE)

#qPCR-derived Mackin scores were generally higher than RFTM scores. Among oysters classified as uninfected by RFTM, 17.0% were qPCR-positive, with 69.7% of these qPCR-positive samples assigned low Mackin scores (0.5–1). This pattern may reflect greater detection of low-level infections by qPCR, although differences in tissue analyzed between methods may also contribute to the observed discrepancy.
```
#let's see if there is any statistical differences between the methods (rftm and qPCR)
```{r}

#non-parametric wilcoxon sum-rank test 
# Test for significant differences between paired diagnostic methods
wilcox_result <- wilcox.test(rftm_v_qpcr$rftm_final, 
                             rftm_v_qpcr$mackin_score, 
                             paired = TRUE)

print(wilcox_result)


#Regression using mixed effects model
library(lme4)
library(tidyr)
library(dplyr)

#Reshape data to long format so each sample has two rows (one per method)
long_data <- rftm_v_qpcr %>%
  select(Sample_ID, rftm_final, mackin_score) %>% # Replace sample_id with your actual unique identifier column
  pivot_longer(cols = c(rftm_final, mackin_score), 
               names_to = "Method", 
               values_to = "Score")
# Install and load the ordinal package if you haven't already
# install.packages("ordinal")
library(ordinal)

# 1. Convert your Score column into a strictly ordered factor
long_data$Score <- factor(long_data$Score, 
                          levels = c(0, 0.5, 1, 2, 3, 4, 5), 
                          ordered = TRUE)

# 2. Run a Cumulative Link Mixed Model with Sample_ID as a random effect
ordinal_paired_model <- clmm(Score ~ Method + (1 | Sample_ID), 
                             data = long_data)

# 3. View the stable summary metrics
summary(ordinal_paired_model)

rftm_v_qpcr %>%
  summarise(
    median_RFTM = median(rftm_final, na.rm = TRUE),
    median_Mackin = median(mackin_score, na.rm = TRUE),
    median_difference = median(rftm_final - mackin_score, na.rm = TRUE),
    mean_difference = mean(rftm_final - mackin_score, na.rm = TRUE)
  )

rftm_v_qpcr %>%
  mutate(difference = rftm_final - mackin_score) %>%
  summarise(
    n = sum(!is.na(difference)),
    median_difference = median(difference, na.rm = TRUE),
    mean_difference = mean(difference, na.rm = TRUE),
    min_difference = min(difference, na.rm = TRUE),
    max_difference = max(difference, na.rm = TRUE)
  )

agreement <- rftm_v_qpcr %>%
  dplyr::filter(
    !is.na(rftm_final),
    !is.na(mackin_score)
  ) %>%
  dplyr::mutate(
    difference = rftm_final - mackin_score,
    exact_agreement = rftm_final == mackin_score,
    within_1 = abs(difference) <= 1
  ) %>%
  dplyr::summarise(
    n = length(rftm_final),
    exact_agreement_pct = mean(exact_agreement) * 100,
    within_1_pct = mean(within_1) * 100
  )

agreement

#RFTM tends to give lower scores than qPCR (significant difference), however, overall we are getting similar results. 33.5% had exactly the same score and 79.5% were within 1 disease intensity category and typically the difference was only 0.5 with lower RFTM. 

#"RFTM and qPCR-derived Mackin disease-intensity scores differed significantly (paired Wilcoxon signed-rank test, V = 3958, p < 0.001). RFTM scores were generally lower than qPCR-derived Mackin scores, with a median paired difference of −0.5 score categories (RFTM − Mackin). Exact agreement between methods occurred for 33.5% of samples, while 79.5% of samples differed by no more than one score category."
```

#Lets compare the histology scores to the qPCR. qPCR is more sensitive so we may see slightly inflated scores but I think it is good to check in terms of postive and negative results. What I am looking for is if qPCR is lower than RFTM or is negative and histopathology is positive. 
```{r}
#pull in histopathology data
Histo_all <- read.csv("Master_files/histopathology_scoring_all_data.csv")


#Merge MSX qPCR data

histo_v_qpcr <- merge(Histo_all, MSX_mackin_scores, by = c("Sample_ID", "Site"))
histo_v_qpcr <- histo_v_qpcr %>% select(Sample_ID, Site, haplo, mackin_score , Copies_ul_ng.ul)
histo_v_qpcr

histo_v_qpcr$results = ifelse(histo_v_qpcr$mackin_score < histo_v_qpcr$haplo, 'FALSE',
  ifelse(histo_v_qpcr$mackin_score > histo_v_qpcr$haplo, 'TRUE',
  ifelse(histo_v_qpcr$mackin_score == histo_v_qpcr$haplo, 'TRUE', NA)))
  
histo_v_qpcr

FALSE_histo_qpcr<- histo_v_qpcr %>% filter( results == 'FALSE')%>% mutate(
    rerun_assay = NA_character_,
    rerun_date = as.Date(NA),
    rerun_result = NA_real_,
    rerun_mackin_score = NA_real_,
    rerun_status = "Pending"
  ) #adding columns we can fill in later on once we rerun the samples to identify if these results are true
FALSE_histo_qpcr

write.csv(FALSE_histo_qpcr, "Lab_Data_qPCR\\outputs/qPCR_sample_rerun_list_MSX.csv", row.names=FALSE)
```
#let's see if there is any statistical differences between the methods
```{r}

#non-parametric wilcoxon sum-rank test 
# Test for significant differences between paired diagnostic methods
wilcox_result <- wilcox.test(histo_v_qpcr$haplo, 
                             histo_v_qpcr$mackin_score, 
                             paired = TRUE)

print(wilcox_result)


#Regression using mixed effects model
library(lme4)
library(tidyr)
library(dplyr)

# 1. Reshape data to long format so each sample has two rows (one per method)
long_data <- histo_v_qpcr %>%
  select(Sample_ID, haplo, mackin_score) %>% # Replace sample_id with your actual unique identifier column
  pivot_longer(cols = c(haplo, mackin_score), 
               names_to = "Method", 
               values_to = "Score")

# 2. Fit a mixed model treating sample_id as a random effect to handle pairing
method_model <- glmer(Score ~ Method + (1 | Sample_ID), 
                      data = long_data, 
                      family = poisson(link = "log")) # Use poisson or ordinal/clmm if Score is strict factor

summary(method_model)

#qPCR is reading significantly higher intensity scores than histology

```
