---
title: "qPCR QAQC of results"
author: "Mariah Kachmar"
date: "2025-07-09"
output: html_document
Description: The purpose of this script is to import, clean up, and summarize Triplex qPCR data for the detection of Dermo, MSX, and SSO to determine prevalence and infection intensity for the LISS Oyster Health Project's monthly sampling at Ash Creek and Fence Creek intertidal sites in Connecticut and Goldstar beach and Laurel Hollow subtidal sites on Long Island, NY. Infection intensity will be reported as copies/ng. This data will then be cross validated with RFTM and Histology data. 

Last updated:8/14/26 by Mariah kachmar

- working on inputting files & a QAQC routine

---

#Set working directory
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/mariah.kachmar/documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance') 
#knitr::opts_knit$set(root.dir = 'C:/Users/kyra.lenderman/documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance') 
#knitr::opts_knit$set(root.dir = 'C:/Users/sarina.dery/documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance') 
```

#run packages
```{r}
library("ggplot2")
library("readxl")
library("plyr")                                     
library("dplyr")
library("Rmisc")
library("readr")  
library("lubridate")
library("purrr")
library(reshape2)
library(stringr)
library("RColorBrewer")
library("stringr")
library("tidyverse")
```


#### This code chunk merges all .csv files within the qubit folder into one data fram and outputs the full dataset into a .csv master file. This allows us to download the raw data as a .csv, add it to the repository folder, and create the master data file without copying and pasting data in excel.
# reading in .csv files from local folder
```{r}

data_all <- list.files(path = "Lab_Data_qPCR/raw_data",                           # Identify all CSV files
  pattern = "*.csv", full.names = TRUE) %>% 
  lapply(read_csv, skip= 19) %>%  # Store all files in list
  bind_rows                       # Combine data sets into one data set 
data_all                          # Print data to RStudio console

as.data.frame(data_all)   # Convert tibble to data.frame


data_all <- data_all %>% filter(!is.na('Content')) #removing any NA or blank data

#What columns are necessary? I am removing 'Well' (Well # on the 96 well plate)  

data_all<- data_all %>% dplyr::select(c('Fluor','Content', 'Cq', 'Starting Quantity (SQ)', 'Cq Mean', 'Cq Std. Dev', 'SQ Std. Dev'))

#lets rename these columns to be more user friendly
data_all <- data_all %>% dplyr::rename(Target_dye = 'Fluor',Sample_ID = 'Content', Concentration = 'Starting Quantity (SQ)', St_dev = 'SQ Std. Dev', Cq_st_dev = 'Cq Std. Dev', Cq_mean = 'Cq Mean' )

#We need dates
data_all <- data_all %>%   mutate(
    # extract parts
    Month_year = str_sub(Sample_ID, 1, 4),
    Site = str_extract(Sample_ID, "(?<=\\d{4})[A-Za-z]+(?=_)"),
    Replicate = str_extract(Sample_ID, "(?<=_)\\d+"),
    # convert month-year to real date (first day of month)
    Date = my(Month_year)
  )

data_all

write.csv(data_all, "Master_files\\qPCR_all_data.csv", row.names=FALSE)
```

#DO NOT RUN THIS CHUNK 

#playing around with code to identify when the standards do not amplify (specifically 10^1). This is in early stages - not finished. The above chunk should be used until further instructions. 
```{r}
# 1. Identify all CSV files with their full paths
file_paths <- list.files(
  path = "Lab_Data_qPCR/raw_data", 
  pattern = "*.csv", 
  full.names = TRUE
)

#Name the vector with itself so bind_rows can use the full paths
file_paths <- set_names(file_paths)

# Read, combine, and retain full file paths in the "Source_File" column
data_qpcr <- file_paths %>% 
  map(~ read_csv(.x, skip = 19, na = c("NA", "NaN"))) %>% 
  bind_rows(.id = "Source_File") 

#What columns are necessary? I am removing 'Well' (Well # on the 96 well plate)  

data_qpcr<- data_qpcr %>% dplyr::select(c('Source_File','Fluor','Content', 'Cq', 'Starting Quantity (SQ)', 'Cq Mean', 'Cq Std. Dev', 'SQ Std. Dev'))

#lets rename these columns to be more user friendly
data_qpcr <- data_qpcr %>% dplyr::rename(Target_dye = 'Fluor',Sample_ID = 'Content', Concentration = 'Starting Quantity (SQ)', St_dev = 'SQ Std. Dev', Cq_st_dev = 'Cq Std. Dev', Cq_mean = 'Cq Mean' )

#We need dates
data_qpcr <- data_qpcr %>%   mutate(
    # extract parts
    Month_year = str_sub(Sample_ID, 1, 4),
    Site = str_extract(Sample_ID, "(?<=\\d{4})[A-Za-z]+(?=_)"),
    Replicate = str_extract(Sample_ID, "(?<=_)\\d+"),
    # convert month-year to real date (first day of month)
    Date = my(Month_year)
  )

# Filter for failed standard curve samples
failed_standards <- data_qpcr %>%
  filter(str_detect(Sample_ID, regex("^Std-", ignore_case = TRUE))) %>%
  filter(Cq == 0 | is.na(Cq) | is.nan(Cq)) %>%
  select(Source_File, Sample_ID, Cq, Target_dye, Concentration)

# View 
print(data_qpcr)
print(failed_standards)

```


```{r}
#Percentage of standard deviations >0.5 for qPCR Cq values
high_cq_sd <- data_all %>% filter(Cq_st_dev > 0.5) %>% nrow() #creates number of rows that have sd # >0.5
high_cq_sd_list <- data_all %>% filter(Cq_st_dev > 0.5) #creates list of which samples have high sd
percentage_highsd <- (high_cq_sd/nrow(data_all)) * 100 # determining percentage of samples have high standard deviation
percentage_highsd #outputs percentage


#Percentage of standard deviations >0.5 for qPCR technical replicate values
high_sd <- data_all %>% filter(St_dev > 0.5) %>% nrow() #creates number of rows that have sd # >0.5
high_sd_list <- data_all %>% filter(St_dev > 0.5) #creates list of which samples have high sd
percentage_highsd <- (high_sd/nrow(data_all)) * 100 # determining percentage of samples have high standard deviation
percentage_highsd #outputs percentage

#Identify which samples have a standard deviation > 0.5 for qPCR technical replicate values
high_sd_list <- data_all %>% filter(St_dev >0.05) %>% arrange(desc(St_dev))
high_sd_list

```

```{r}
#Lets create means of our sample duplicates 

data_summary <- data_all %>%
  dplyr::mutate(Concentration = as.numeric(Concentration)) %>%
  dplyr::group_by(Target_dye,Sample_ID, Site, Date) %>%
  dplyr::summarise(
    Copies_per_2ul = mean(Concentration, na.rm = TRUE),
    Copies_per_2ul_sd = sd(Concentration, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

data_summary


#now we need copies per ul for reporting

data_summary <- data_summary %>%
  mutate(
    Copies_per_uL =  Copies_per_2ul / 2,
    Copies_per_uL_sd =  Copies_per_2ul_sd / 2
  )
  
data_summary



```

```{r}

#merge Qubit data

#bring in data
data_quibit <- read.csv("Lab_Data_DNAQuantification/output/qubit_summary.csv")
data_quibit


merged_data <- merge(data_summary, data_quibit, by = "Sample_ID")
merged_data

```

```{r}

#Lets complete those back calculations

data_summary <- merged_data %>%
  mutate(
    Ratio_copies_per_ng =  Copies_per_uL / mean_concentration_ng.uL,
    Copies_ul_ng.ul =  Ratio_copies_per_ng * 33.33, #this will be the final concentration per 1 ul, roger williams uses this concentration (what we should report for scoring?)
    Copies_2ul_final = Copies_ul_ng.ul *2 #this is the final concentration per reaction
  )
  
data_summary

#add a year column 
data_summary <- data_summary %>%
  mutate(Year = year(Date))

```

```{r}

#lets work on some code that bins our copies/ul into the Mackin's scale. This is the same scale that Roger Williams Univ. is using currently and may be subject to change based on our data. This is preliminary to test code!!

#Let's parse our dermo data (Target_dye = FAM)

dermo_qPCR_summary <- data_summary %>% filter(Target_dye == 'FAM')
dermo_qPCR_summary

qpcr_mackin_scores <- dermo_qPCR_summary %>%
  mutate(mackin_score = case_when(
    Copies_ul_ng.ul < 10 ~ 0,
    Copies_ul_ng.ul >= 10 & Copies_ul_ng.ul < 100 ~ 0.5, #light
    Copies_ul_ng.ul >= 100 & Copies_ul_ng.ul < 2000 ~ 1, #light
    Copies_ul_ng.ul >= 2000 & Copies_ul_ng.ul < 10000 ~ 2, #moderate
    Copies_ul_ng.ul >= 10000 & Copies_ul_ng.ul < 100000 ~ 3, #moderate
    Copies_ul_ng.ul >= 100000 & Copies_ul_ng.ul < 1000000 ~ 4, #heavy
    Copies_ul_ng.ul >= 1000000 ~ 5, #heavy
    TRUE ~ 0
  ))

qpcr_mackin_scores

#Lets do the same for MSX (Target_dye = HEX)


MSX_qPCR_summary <- data_summary %>% filter(Target_dye == 'HEX')
MSX_qPCR_summary

MSX_mackin_scores <- MSX_qPCR_summary %>%
  mutate(mackin_score = case_when(
    Copies_ul_ng.ul < 10 ~ 0,
    Copies_ul_ng.ul >= 10 & Copies_ul_ng.ul < 10000 ~ 1,
    Copies_ul_ng.ul >= 10000 & Copies_ul_ng.ul < 1000000 ~ 2,
    Copies_ul_ng.ul >= 1000000  ~ 3,
    TRUE ~ 0
  ))

MSX_mackin_scores

```

```{r}

#Now for some prevalence and intensity calculations

#weighted prevalence 

#dermo
d_wp_summary <- qpcr_mackin_scores %>%
  dplyr::group_by(Site, Date, Year) %>% 
  dplyr::summarize( Weighted_prevalence = mean(mackin_score))
d_wp_summary

#dermo graph
ggplot(data=d_wp_summary, aes(x=Date, y=Weighted_prevalence, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Weighted Prevalence of Dermo infection (qPCR)", x ="Date", y = "Weighted Prevalence Score")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "fixed", ncol=1,)+ scale_x_date(
  date_breaks = "1 month",
  date_labels = "%b/%Y")#scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))

#MSX
msx_wp_summary<- MSX_mackin_scores %>%
  dplyr::group_by(Site, Date, Year) %>% 
  dplyr::summarize( Weighted_prevalence = mean(mackin_score))

msx_wp_summary

#MSX graph

ggplot(data=msx_wp_summary, aes(x=Date, y=Weighted_prevalence, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Weighted Prevalence of MSX infection (qPCR)", x ="Date", y = "Weighted Prevalence Score")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "fixed", ncol=1,)+ scale_x_date(
  date_breaks = "1 month",
  date_labels = "%b/%Y")#scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))

```

# Prevalence calculations for all sites
# Prevalence = # infected/ total # sampled
```{r}

## Prevalence T/F for DERMO
qpcr_mackin_scores<-qpcr_mackin_scores%>% filter(!is.na(mackin_score))
qpcr_mackin_scores

qpcr_mackin_scores$prevalent = ifelse(qpcr_mackin_scores$mackin_score == 0, '0',
  ifelse(qpcr_mackin_scores$mackin_score > 0, '1', NA))
  
qpcr_mackin_scores

# YEARLY PREVALENCE %
present <- nrow(qpcr_mackin_scores %>% filter(prevalent == '1'))
absent <- nrow(qpcr_mackin_scores %>% filter(prevalent == '0'))

prevalence_all <- present/(present+absent)
prevalence_all


str(qpcr_mackin_scores$prevalent)
qpcr_mackin_scores$prevalent <- as.numeric(qpcr_mackin_scores$prevalent)

d_prevalence_summary <- qpcr_mackin_scores %>%
  dplyr::group_by(Site, Year, Date) %>% 
  dplyr::summarize( prevalence_precentage= mean(prevalent)* 100)
d_prevalence_summary


#Graph DERMO

ggplot(data=d_prevalence_summary, aes(x=Date, y=prevalence_precentage, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Prevalence of Dermo infection (qPCR)", x ="Date", y = "Dermo Prevlaence %")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "free", ncol=1) + scale_x_date(
  date_breaks = "1 month",
  date_labels = "%b/%Y")#+ scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))


## Prevalence T/F for MSX

MSX_mackin_scores<-MSX_mackin_scores%>% filter(!is.na(mackin_score))
MSX_mackin_scores

MSX_mackin_scores$prevalent = ifelse(MSX_mackin_scores$mackin_score == 0, '0',
  ifelse(MSX_mackin_scores$mackin_score > 0, '1', NA))
  
MSX_mackin_scores

# YEARLY PREVALENCE %
present <- nrow(MSX_mackin_scores %>% filter(prevalent == '1'))
absent <- nrow(MSX_mackin_scores %>% filter(prevalent == '0'))

prevalence_all <- present/(present+absent)
prevalence_all


str(MSX_mackin_scores$prevalent)
MSX_mackin_scores$prevalent <- as.numeric(MSX_mackin_scores$prevalent)

MSX_prevalence_summary <- MSX_mackin_scores %>%
  dplyr::group_by(Site, Year, Date) %>% 
  dplyr::summarize( prevalence_precentage= mean(prevalent)* 100)
MSX_prevalence_summary

#Graph MSX

ggplot(data=MSX_prevalence_summary, aes(x=Date, y=prevalence_precentage, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Prevalence of MSX infection (qPCR)", x ="Date", y = "MSX Prevalence %")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "free", ncol=1)+ scale_x_date(
  date_breaks = "1 month",
  date_labels = "%b/%Y")#scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))


```

#Lets compare the RFTM scores to the qPCR. qPCR is more sensitive so we may see slightly inflated scores but I think it is good to check in terms of postive and negative results. RFTM we are also only taking mantle tissue vs qPCR we are taking GGR. What I am looking for is if qPCR is lower than RFTM or is negative and RFTM is positive. 
```{r}
#pull in RFTM data
RFTM_all <- read.csv("Master_files/RFTM_all_data.csv")
RFTM_all <- RFTM_all %>% dplyr::rename("Sample_ID" = "lab_id")

#Merge dermo qPCR data

rftm_v_qpcr <- merge(RFTM_all, qpcr_mackin_scores, by = "Sample_ID")
rftm_v_qpcr <- rftm_v_qpcr %>% select(Sample_ID, site, rftm_final, mackin_score , Copies_ul_ng.ul)
rftm_v_qpcr

rftm_v_qpcr$results = ifelse(rftm_v_qpcr$mackin_score < rftm_v_qpcr$rftm_final, 'FALSE',
  ifelse(rftm_v_qpcr$mackin_score > rftm_v_qpcr$rftm_final, 'TRUE',
  ifelse(rftm_v_qpcr$mackin_score == rftm_v_qpcr$rftm_final, 'TRUE', NA)))
  
rftm_v_qpcr

FALSE_qpcr<- rftm_v_qpcr %>% filter( results == 'FALSE') %>%
  mutate(
    rerun_assay = NA_character_,
    rerun_date = as.Date(NA),
    rerun_result = NA_real_,
    rerun_mackin_score = NA_real_,
    rerun_status = "Pending"
  ) #adding columns we can fill in later on once we rerun the samples to identify if these results are true
FALSE_qpcr

write.csv(FALSE_qpcr, "Lab_Data_qPCR\\outputs/qPCR_sample_rerun_list.csv", row.names=FALSE)
```

#Lets compare the histology scores to the qPCR. qPCR is more sensitive so we may see slightly inflated scores but I think it is good to check in terms of postive and negative results. What I am looking for is if qPCR is lower than RFTM or is negative and histopathology is positive. 
```{r}
#pull in histopathology data
Histo_all <- read.csv("Master_files/histopathology_scoring_all_data.csv")


#Merge MSX qPCR data

histo_v_qpcr <- merge(Histo_all, MSX_mackin_scores, by = c("Sample_ID", "Site"))
histo_v_qpcr <- histo_v_qpcr %>% select(Sample_ID, Site, haplo, mackin_score , Copies_ul_ng.ul)
histo_v_qpcr

histo_v_qpcr$results = ifelse(histo_v_qpcr$mackin_score < histo_v_qpcr$haplo, 'FALSE',
  ifelse(histo_v_qpcr$mackin_score > histo_v_qpcr$haplo, 'TRUE',
  ifelse(histo_v_qpcr$mackin_score == histo_v_qpcr$haplo, 'TRUE', NA)))
  
histo_v_qpcr

FALSE_histo_qpcr<- histo_v_qpcr %>% filter( results == 'FALSE')%>% mutate(
    rerun_assay = NA_character_,
    rerun_date = as.Date(NA),
    rerun_result = NA_real_,
    rerun_mackin_score = NA_real_,
    rerun_status = "Pending"
  ) #adding columns we can fill in later on once we rerun the samples to identify if these results are true
FALSE_histo_qpcr

write.csv(FALSE_histo_qpcr, "Lab_Data_qPCR\\outputs/qPCR_sample_rerun_list_MSX.csv", row.names=FALSE)
```