---
title: "qPCR QAQC of results"
author: "Mariah Kachmar"
date: "2025-07-09"
output: html_document
Description: The purpose of this script is to import, clean up, and summarize Triplex qPCR data for the detection of Dermo, MSX, and SSO to determine prevalence and infection intensity for the LISS Oyster Health Project's monthly sampling at Ash Creek and Fence Creek intertidal sites in Connecticut and Goldstar beach and Laurel Hollow subtidal sites on Long Island, NY. Infection intensity will be reported as copies/ng. This data will then be cross validated with RFTM and Histology data. 

Last updated:5/7/26 by Mariah kachmar

- working on inputting files & a QAQC routine
- got to calculating copies per ul 
- next steps are the back calculations to standardize the samples - ratio copies per ng, copies/ul in 50ng/ul, then copies/2ul based on 50ng/ul
---

#Set working directory
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/mariah.kachmar/documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance') #Mariah's
#knitr::opts_knit$set(root.dir = 'C:/Users/kyra.lenderman/documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance') #Kyra's
#knitr::opts_knit$set(root.dir = 'C:/Users/sarina.dery/documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance') #Kelly's
```

#run packages
```{r}
library("ggplot2")
library("readxl")
library("plyr")                                     
library("dplyr")
library("Rmisc")
library("readr")  
library("lubridate")
library("purrr")
library(reshape2)
library(stringr)
library("RColorBrewer")
library("stringr")
library("tidyverse")
```

#### This code chunk merges all .csv files within the qubit folder into one data fram and outputs the full dataset into a .csv master file. This allows us to download the raw data as a .csv, add it to the repository folder, and create the master data file without copying and pasting data in excel.
# reading in .csv files from local folder
```{r}



data_all <- list.files(path = "Lab_Data_qPCR/raw_data",                           # Identify all CSV files
  pattern = "*.csv", full.names = TRUE) %>% 
  lapply(read_csv, skip= 19) %>%                              # Store all files in list
  bind_rows                                         # Combine data sets into one data set 
data_all                                            # Print data to RStudio console

as.data.frame(data_all)                            # Convert tibble to data.frame


data_all <- data_all %>% filter(!is.na('Content'))

#What columns are necessary? I am removing 'Well' (Well # on the 96 well plate)  

data_all<- data_all %>% select(c('Fluor','Content', 'Cq', 'Starting Quantity (SQ)', 'Cq Mean', 'Cq Std. Dev', 'SQ Std. Dev'))

#lets rename these columns to be more user friendly
data_all <- data_all %>% dplyr::rename(Target_dye = 'Fluor',Sample_ID = 'Content', Concentration = 'Starting Quantity (SQ)', St_dev = 'SQ Std. Dev', Cq_st_dev = 'Cq Std. Dev', Cq_mean = 'Cq Mean' )


data_all <- data_all %>%   mutate(
    # extract parts
    Month_year = str_sub(Sample_ID, 1, 4),
    Site = str_extract(Sample_ID, "(?<=\\d{4})[A-Za-z]+(?=_)"),
    Replicate = str_extract(Sample_ID, "(?<=_)\\d+"),
    # convert month-year to real date (first day of month)
    Date = my(Month_year)
  )

data_all

write.csv(data_all, "Master_files\\qPCR_all_data.csv", row.names=FALSE)
```

```{r}
#Percentage of standard deviations >0.5 for qPCR Cq values
high_cq_sd <- data_all %>% filter(Cq_st_dev > 0.5) %>% nrow() #creates number of rows that have sd # >0.5
high_cq_sd_list <- data_all %>% filter(Cq_st_dev > 0.5) #creates list of which samples have high sd
percentage_highsd <- (high_cq_sd/nrow(data_all)) * 100 # determining percentage of samples have high standard deviation
percentage_highsd #outputs percentage

```

```{r}
#Lets create means of our sample duplicates 

data_summary <- data_all %>%
  dplyr::mutate(Concentration = as.numeric(Concentration)) %>%
  dplyr::group_by(Target_dye,Sample_ID, Site, Date) %>%
  dplyr::summarise(
    Copies_per_2ul = mean(Concentration, na.rm = TRUE),
    Copies_per_2ul_sd = sd(Concentration, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

data_summary


#now we need copies per ul for reporting

data_summary <- data_summary %>%
  mutate(
    Copies_per_uL =  Copies_per_2ul / 2,
    Copies_per_uL_sd =  Copies_per_2ul_sd / 2
  )
  
data_summary



```

```{r}

#merge Qubit data

#bring in data
data_quibit <- read.csv("Lab_Data_DNAQuantification/output/qubit_summary.csv")
data_quibit


merged_data <- merge(data_summary, data_quibit, by = "Sample_ID")
merged_data

```

```{r}

#Lets complete those back calculations

data_summary <- merged_data %>%
  mutate(
    Ratio_copies_per_ng =  Copies_per_uL / mean_concentration_ng.uL,
    Copies_ul_ng.ul =  Ratio_copies_per_ng * 33.33, #this will be the final concentration per 1 ul (what we should report for scoring?)
    Copies_2ul_final = Copies_ul_ng.ul *2 #this is the final concentration per reaction
  )
  
data_summary

#add a year column 
data_summary <- data_summary %>%
  mutate(Year = year(Date))

```

```{r}

#lets work on some code that bins our copies/ul into the Mackin's scale. This is the same scale that Roger Williams Univ. is using currently and may be subject to change based on our data. This is preliminary to test code!!

#Let's parse our dermo data (Target_dye = FAM)

dermo_qPCR_summary <- data_summary %>% filter(Target_dye == 'FAM')
dermo_qPCR_summary

qpcr_mackin_scores <- dermo_qPCR_summary %>%
  mutate(mackin_score = case_when(
    Copies_ul_ng.ul < 10 ~ 0,
    Copies_ul_ng.ul >= 10 & Copies_ul_ng.ul < 100 ~ 0.5,
    Copies_ul_ng.ul >= 100 & Copies_ul_ng.ul < 2000 ~ 1,
    Copies_ul_ng.ul >= 2000 & Copies_ul_ng.ul < 10000 ~ 2,
    Copies_ul_ng.ul >= 10000 & Copies_ul_ng.ul < 100000 ~ 3,
    Copies_ul_ng.ul >= 100000 & Copies_ul_ng.ul < 1000000 ~ 4,
    Copies_ul_ng.ul >= 1000000 ~ 5,
    TRUE ~ 0
  ))

qpcr_mackin_scores

#Lets do the same for MSX (Target_dye = HEX)


MSX_qPCR_summary <- data_summary %>% filter(Target_dye == 'HEX')
MSX_qPCR_summary

MSX_mackin_scores <- MSX_qPCR_summary %>%
  mutate(mackin_score = case_when(
    Copies_ul_ng.ul < 10 ~ 0,
    Copies_ul_ng.ul >= 10 & Copies_ul_ng.ul < 10000 ~ 1,
    Copies_ul_ng.ul >= 10000 & Copies_ul_ng.ul < 1000000 ~ 2,
    Copies_ul_ng.ul >= 1000000  ~ 3,
    TRUE ~ 0
  ))

MSX_mackin_scores

```

```{r}

#Now for some prevalence and intensity calculations

#weighted prevalence 

#dermo
d_wp_summary <- qpcr_mackin_scores %>%
  dplyr::group_by(Site, Date, Year) %>% 
  dplyr::summarize( Weighted_prevalence = mean(mackin_score))
d_wp_summary

#dermo graph
ggplot(data=d_wp_summary, aes(x=Date, y=Weighted_prevalence, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Weighted Prevalence of Dermo infection (qPCR)", x ="Date", y = "Weighted Prevalence Score")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "fixed", ncol=1,)+scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))

#MSX
msx_wp_summary<- MSX_mackin_scores %>%
  dplyr::group_by(Site, Date, Year) %>% 
  dplyr::summarize( Weighted_prevalence = mean(mackin_score))

msx_wp_summary

#MSX graph

ggplot(data=msx_wp_summary, aes(x=Date, y=Weighted_prevalence, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Weighted Prevalence of MSX infection (qPCR)", x ="Date", y = "Weighted Prevalence Score")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "fixed", ncol=1,)+scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))

```

# Prevalence calculations for all sites
# Prevalence = # infected/ total # sampled
```{r}

## Prevalence T/F for DERMO
qpcr_mackin_scores<-qpcr_mackin_scores%>% filter(!is.na(mackin_score))
qpcr_mackin_scores

qpcr_mackin_scores$prevalent = ifelse(qpcr_mackin_scores$mackin_score == 0, '0',
  ifelse(qpcr_mackin_scores$mackin_score > 0, '1', NA))
  
qpcr_mackin_scores

# YEARLY PREVALENCE %
present <- nrow(qpcr_mackin_scores %>% filter(prevalent == '1'))
absent <- nrow(qpcr_mackin_scores %>% filter(prevalent == '0'))

prevalence_all <- present/(present+absent)
prevalence_all


str(qpcr_mackin_scores$prevalent)
qpcr_mackin_scores$prevalent <- as.numeric(qpcr_mackin_scores$prevalent)

d_prevalence_summary <- qpcr_mackin_scores %>%
  dplyr::group_by(Site, Year, Date) %>% 
  dplyr::summarize( prevalence_precentage= mean(prevalent)* 100)
d_prevalence_summary


#Graph DERMO

ggplot(data=d_prevalence_summary, aes(x=Date, y=prevalence_precentage, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Prevalence of Dermo infection (qPCR)", x ="Date", y = "Dermo Prevlaence %")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "free", ncol=1)+scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))


## Prevalence T/F for MSX

MSX_mackin_scores<-MSX_mackin_scores%>% filter(!is.na(mackin_score))
MSX_mackin_scores

MSX_mackin_scores$prevalent = ifelse(MSX_mackin_scores$mackin_score == 0, '0',
  ifelse(MSX_mackin_scores$mackin_score > 0, '1', NA))
  
MSX_mackin_scores

# YEARLY PREVALENCE %
present <- nrow(MSX_mackin_scores %>% filter(prevalent == '1'))
absent <- nrow(MSX_mackin_scores %>% filter(prevalent == '0'))

prevalence_all <- present/(present+absent)
prevalence_all


str(MSX_mackin_scores$prevalent)
MSX_mackin_scores$prevalent <- as.numeric(MSX_mackin_scores$prevalent)

MSX_prevalence_summary <- MSX_mackin_scores %>%
  dplyr::group_by(Site, Year, Date) %>% 
  dplyr::summarize( prevalence_precentage= mean(prevalent)* 100)
MSX_prevalence_summary

#Graph MSX

ggplot(data=MSX_prevalence_summary, aes(x=Date, y=prevalence_precentage, color=Site)) +
  geom_point()+  
  geom_line(aes(group = Site))+
   theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title="Prevalence of MSX infection (qPCR)", x ="Date", y = "MSX Prevalence %")+ theme(axis.title.y = element_text(size = rel(1.3), angle =90), axis.title.x = element_text(size = rel(1.3), angle = 0))+
  theme(axis.text=element_text(size=12)) + facet_wrap(~ Year, scales = "free", ncol=1)+scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))


```