---
title: "Sonde Preliminary Statistics"
author: "Mariah Kachmar"
date: "2026-01-26"
output: html_document

description: This code is to conduct preliminary statistics on the sonde data to see general trends within sites and between sites. 
---

#How to deal with winter months that have no data? or data gaps in general?
#Combine Hobo & sonde data to fill in gaps for temperature?
#find data to input?
#Seasonal Imputation? 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/mariah.kachmar/Documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance/Sonde_Data') #Mariah's
#knitr::opts_knit$set(root.dir = 'C:/Users/kyra.lenderman/Documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance/Sonde_Data') #Kyra's
```

```{r,echo = FALSE}
#this code loads all packages that the code will need
#make sure all packages are installed before running this code
library("ggplot2")
library("ggpubr")
library("readxl")
library("plyr")                                     
library("dplyr")
library("Rmisc")
library("readr")  
library("lubridate")
library("purrr")
library("tidyr")
library("survminer")
library("survival")
library("ggpubr")
library(glmm)
library(glmmTMB)
library(effectsize)
library("mgcv")
```

## Load data:

```{r load data}

Master_table <- read.csv("output/Sonde_master.csv",check.names=FALSE) %>% dplyr::select(-1) # check.names = F allows with col names with spaces (opthwise fills with a '.' )

Master_table$Date.Time <- lubridate::ymd_hms(Master_table$Date.Time)

```

```{r}
#melting the data to be long format
# lets create monthly means for statistical analysis
library(dplyr)
library(lubridate)
library(reshape2)
library(data.table)

Sonde_monthly_means_LONG <-
  Master_table %>%
    # convert Date.Time to a proper POSIXct if not already
    mutate(Date.Time = ymd_hms(Date.Time)) %>%
    # round to month (first day of month)
    mutate(
      Year = year(Date.Time),
      Month = floor_date(Date.Time, unit = "month")) %>%
    # reshape to long format
    reshape2::melt(
      id.vars = c("Year", "Month", "Site"),
      measure.vars = c("Actual.Conductivity", "Salinity", "Resistivity",
                       "Total.Dissolved.Solids", "RDO.Concentration",
                       "RDO.Saturation", "Oxygen.Partial.Pressure",
                       "Chlorophyll.a.Fluorescence", "pH", "pH.mV", "Temperature")
    ) %>%
    # group by month, site, and variable
    dplyr::group_by(Year, Month, Site, variable) %>%
    # calculate mean, sd, and se for each group
    dplyr::summarise(
      mean = mean(value, na.rm = TRUE),
      sd   = sd(value, na.rm = TRUE),
      se   = sd(value, na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )

vars <- as.data.frame(
          list(
            as.character(
              unique(
                Sonde_monthly_means_LONG$variable)))) %>% 
        dplyr::rename(dat = 1)

write.csv(Sonde_monthly_means_LONG, "output/Sonde_monthly_means_master.csv")

## We don't want this in long format for all analysis - especially when we merge with other data (disease, body condition, etc.)
#we still want means
Sonde_monthly_means_wide <- Sonde_monthly_means_LONG %>% pivot_wider(names_from = variable, values_from = c("mean", "sd", "se"))
write.csv(Sonde_monthly_means_wide, "output/Sonde_monthly_means_wide_master.csv")


```

#Statistical Analysis
#gams are a type of glm that allows for flexible, non-linear relationships between preditor variables and the response variable
#https://m-clark.github.io/generalized-additive-models/application.html
#https://cran.r-project.org/web/packages/gam/gam.pdf
#https://campus.datacamp.com/courses/nonlinear-modeling-with-generalized-additive-models-gams-in-r/interpreting-and-visualizing-gams?ex=1
#How to read summary? adj R^2 provides a measurement of the model fit by penalizing the addition of unnecesssary predictors, indicating how well the model explains the variance; high = capturing relationships effectively , low = not capturing relationships effectively, negative = poor fit
#R.sq >0.7 very good, 0.5-0.7 moderate (acceptable), <0.5 Weak (model doesnt explain much variation); Biological/ecological data: R² ~0.5 is often good due to natural variability.


#Analysis for comparison between CT sites
```{r}
pH_mm_CT         <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'pH' & Site  %in% c('FENC', 'ASHC'))
Salinity_mm_CT     <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'Salinity' & Site %in% c('FENC', 'ASHC'))
Temperature_mm_CT  <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'Temperature' & Site %in% c('FENC', 'ASHC'))
DO_mm_CT         <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'RDO.Concentration' & Site %in% c('FENC', 'ASHC'))
CHLA_mm_CT        <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'Chlorophyll.a.Fluorescence' & Site %in% c('FENC', 'ASHC'))

gam_p_CT <- gam(mean ~ Site + Month + Year , data = pH_mm_CT ) 
summary(gam_p_CT)

gam_S_CT<- gam(mean ~ Site + Month + Year , data = Salinity_mm_CT ) 
summary(gam_S_CT)

gam_T_CT <- gam(mean ~ Site + Month + Year , data = Temperature_mm_CT ) 
summary(gam_T_CT)

gam_D_CT <- gam(mean ~ Site + Month + Year , data = DO_mm_CT ) 
summary(gam_D_CT)

gam_C_CT <- gam(mean ~ Site + Month + Year , data = CHLA_mm_CT ) 
summary(gam_C_CT)

```


#Analysis for Ash creek 
```{r}
pH_mm_ASHC         <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'pH' & Site  %in% c( 'ASHC'))
Salinity_mm_ASHC    <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'Salinity' & Site %in% c('ASHC'))
Temperature_mm_ASHC <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'Temperature' & Site %in% c('ASHC'))
DO_mm_ASHC        <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'RDO.Concentration' & Site %in% c('ASHC'))
CHLA_mm_ASHC       <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'Chlorophyll.a.Fluorescence' & Site %in%c('ASHC'))


gam_p_A <- gam(mean ~ Month + Year , data = pH_mm_ASHC ) 
summary(gam_p_A)

gam_S_A <- gam(mean ~ Month + Year , data = Salinity_mm_ASHC ) 
summary(gam_S_A)

gam_T_A <- gam(mean ~ Month + Year , data = Temperature_mm_ASHC ) 
summary(gam_T_A)

gam_D_A <- gam(mean ~ Month + Year , data = DO_mm_ASHC ) 
summary(gam_D_A)

gam_C_A <- gam(mean ~ Month + Year , data = CHLA_mm_ASHC ) 
summary(gam_C_A)

```

#Analysis for Fence creek 
```{r}
pH_mm_FENC         <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'pH' & Site  %in% c('FENC'))
Salinity_mm_FENC    <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'Salinity' & Site %in% c('FENC'))
Temperature_mm_FENC <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'Temperature' & Site %in% c('FENC'))
DO_mm_FENC        <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'RDO.Concentration' & Site %in% c('FENC'))
CHLA_mm_FENC       <- Sonde_monthly_means_LONG %>% dplyr::filter(variable == 'Chlorophyll.a.Fluorescence' & Site %in%c('FENC'))

gam_p_F <- gam(mean ~ Month + Year , data = pH_mm_FENC ) 
summary(gam_p_F)

gam_S_F <- gam(mean ~ Month + Year , data = Salinity_mm_FENC ) 
summary(gam_S_F)

gam_T_F <- gam(mean ~ Month + Year , data = Temperature_mm_FENC ) 
summary(gam_T_F)

gam_D_F <- gam(mean ~ Month + Year , data = DO_mm_FENC ) 
summary(gam_D_F)

gam_C_F <- gam(mean ~ Month + Year , data = CHLA_mm_FENC ) 
summary(gam_C_F)
```

#Analysis for NY - Gold Star beach
```{r}
```


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
