---
title: "Health and Environment Preliminary Statistics"
author: "Mariah Kachmar"
date: "2026-01-22"
output: html_document

description: This code is for preliminary statistical analysis for all data (environmental, lab, field) comparisons for the EPA Long Island Sound Partnership Oyster Health project. 
---
## Setup: 

* call your root directory, using opts_knit assigns for the entire script so your paths will alwyas be from this start point!

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      cache = TRUE)

#knitr::opts_knit$set(root.dir = "C:/Users/kyra.lenderman/documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance")
knitr::opts_knit$set(root.dir = "C:/Users/mariah.kachmar/documents/Github/EAD-ASEB_EPA_LISS_Disease_Surveillance")

```

#Packaged
```{r}
library("ggplot2")
library("ggpubr")
library("readxl")
library("plyr")                                     
library("dplyr")
library("Rmisc")
library("readr")  
library("lubridate")
library("purrr")
library("tidyr")
library("survminer")
library("survival")
library("ggpubr")
library(glmm)
library(glmmTMB)
library(effectsize)
library("mgcv")
library(car)
library(ppcor)
```

## Load data:
#These are our master files for observed health and environmental parameters
```{r load data}

#Pull in sonde data
Sonde_monthly_means <- read.csv("Sonde_data/output/Sonde_monthly_means_wide_master.csv",check.names=TRUE) %>% dplyr::filter(!Site %in% "LAUR") # check.names = F allows with col names with spaces (opthwise fills with a '.' ) , Also removing site LAUR

#Pull in RFTM data
rftm_summary <- read.csv("Lab_Data_RFTM/output/Completeness_RFTM_intensity_data.csv")

#Pull in body condition data

```
#reconfigure data to be able to merge RFTM & sonde datafiles 
```{r}

#add month and year column to RFTM data
rftm_summary <- rftm_summary %>% dplyr::mutate(Month = month(date_collected), Year = year(date_collected)) %>% rename(Site = site)

#renaming Month column that is actually dates to read correctly and then adding a month column
Sonde_monthly_means <- Sonde_monthly_means %>% rename(date = Month) %>% dplyr::mutate(Month = month(date)) 
                                                             
#merge environmental data and RFTM data by Month, Year, and site
rftm_env_master <- Sonde_monthly_means %>% dplyr::inner_join(rftm_summary, by = c("Month", "Year", "Site"))
rftm_env_master
```


#okay so lets test some stats and see what might be affecting RFTM infection intensity scores (weighted prevalence). 
```{r}
#Lets just start off by messing around with temperature and salinity because we know that they impact the disease intensity


glm_rftm <- glmmTMB(rftm_final ~ mean_Temperature + mean_Salinity + Site + Month + Year + mean_Temperature:Site + mean_Salinity:Site, data = rftm_env_master)
#Anova(glm_ph_rftm)
summary(glm_rftm)

#should we do spearman rank correlations to test relationships? 

# lets separate by site first

# ASH CREEK

AsHC_rftm_env<- rftm_env_master %>%
  filter(Site == "ASHC")

cor.test(AsHC_rftm_env$mean_Temperature,AsHC_rftm_env$rftm_final, method = "spearman" )
#significant positive relationship with temperature (high temps associated with higher disease)
cor.test(AsHC_rftm_env$mean_Salinity,AsHC_rftm_env$rftm_final, method = "spearman" )
#significant positive relationship with salinity (higher salinity associated with higher disease)

#so with both high temp and high salinity may favor increased infections

#Lets try other parameters
cor.test(AsHC_rftm_env$mean_pH,AsHC_rftm_env$rftm_final, method = "spearman" )
cor.test(AsHC_rftm_env$mean_RDO.Concentration,AsHC_rftm_env$rftm_final, method = "spearman" )


#lets look at all three together "is temperature still related to intensity after accounting for salinity? 
pcor.test(
  AsHC_rftm_env$mean_Temperature,
  AsHC_rftm_env$rftm_final,
  AsHC_rftm_env$mean_Salinity,
  method = "spearman"
)

# significant positive erlationship - as one increases the others do

#Fence Creek
FENC_rftm_env<- rftm_env_master %>%
  filter(Site == "FENC")
cor.test(FENC_rftm_env$mean_Temperature,FENC_rftm_env$rftm_final, method = "spearman" )
#not significant - could be other environmental drivers & greater environmental variability
cor.test(FENC_rftm_env$mean_Salinity,FENC_rftm_env$rftm_final, method = "spearman" )
#significant positive relationship with salinity (higher salinity associated with higher disease)

#so salinity might be a more sonsistent predictor of dermo intensity while temperature affects may be more site-specific or influenced by additional factor

#Lets try other parameters
cor.test(FENC_rftm_env$mean_pH,FENC_rftm_env$rftm_final, method = "spearman" ) #not correlated
cor.test(FENC_rftm_env$mean_RDO.Concentration,FENC_rftm_env$rftm_final, method = "spearman" )

#lets look at all three together "is temperature still related to intensity after accounting for salinity? 
pcor.test(
  FENC_rftm_env$mean_Temperature,
  FENC_rftm_env$rftm_final,
  FENC_rftm_env$mean_Salinity,
  method = "spearman"
)
#not significant

#GOLD STAR

GOLD_rftm_env<- rftm_env_master %>%
  filter(Site == "GOLD")
cor.test(GOLD_rftm_env$mean_Temperature,GOLD_rftm_env$rftm_final, method = "spearman",  exact = FALSE )
#not significant - could be other environmental drivers & greater environmental variability
cor.test(GOLD_rftm_env$mean_Salinity,GOLD_rftm_env$rftm_final, method = "spearman",  exact = FALSE )
#moderate positive relationship with salinity, however not statistically significant (higher salinity associated with higher disease)

#so salinity might be a more consistent predictor of dermo intensity while temperature affects may be more site-specific or influenced by additional factor

#Lets try other parameters
cor.test(GOLD_rftm_env$mean_pH,GOLD_rftm_env$rftm_final, method = "spearman",  exact = FALSE ) #not correlated
cor.test(GOLD_rftm_env$mean_RDO.Concentration,GOLD_rftm_env$rftm_final, method = "spearman",  exact = FALSE )

#lets look at all three together "is temperature still related to intensity after accounting for salinity? 
pcor.test(
  GOLD_rftm_env$mean_Temperature,
  GOLD_rftm_env$rftm_final,
  GOLD_rftm_env$mean_Salinity,
  method = "spearman"
)
#no significance or relationship

#so what is driving disease? age? Maybe we can look at the size collections over time in relation to infection intensity

```